---
title: "mc_kienzle"
author: "Thomas M. Rudolf"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(ggplot2)
library(MASS)
library(patchwork)
library(R2OpenBUGS)
#library(rjags)

library(patchwork)
library(tidyverse)
```

Kienzle Model for cutting torque:

$F_{ci} = k_c b h^{1-m_c} = a_p f_z^{1-m_c} sin(\kappa)^{m_c} k_{c1.1} sin(\phi + (k-1) \frac{360}{z})^{1-m_c}$

$M_c = \sum_1^z{F_{ci} r_{tool}}$



$a_p$ cutting depth

$f_z$ feed per cutting edge

$\kappa$ main tool angle

$z$ number of cutting edges

$k$ cutting edgre id : 1..z

The value of $m_c$ and $k_{c1.1}$ are material specific values and have to be identified. A mean value 0.25. One will take a #beta distribution# to simulate the resulting torque for different machining situations. The machining are varied in cutting depth $a_p$, feed per tooth $f_z$ and different entrance angles $\phi$.

```{r}
cat("model{
for (i in 1:k){
Mc[i] = 0
  for(k in 1:z){
    Mc[i] = Mc[i] + (ap * fz^(1-mc) * sin(kappa)^mc * kc1 * sin(phi[i] + (k-1)*2*pi/z)^(1-mc))*rtool
  }
}

mc ~ dbeta(alpha,beta)
alpha <- dnorm(m_alpha, sd_alpha)
beta <- dnorm(m_beta, sd_beta)
  } ", file="jags_kienzle.txt")


ap = 1.0  # cutting depth in mm
fz = 0.15 # feed per tooth in mm
kappa = 5/180*pi # main angle of tool insert in degree
z = 4     # number of cutting edges [2, 3, 4]
kc1 = 1e3 # material value
rtool = 40/2# tool radius in mm
m_alpha = 0.0 
sd_alpha = 0.01
m_beta = 0.0 
sd_beta = 0.01

##

```
